### 2019-01-26 assignments/2/assignment_2.Rmd
Initial commit

---
title: 'EEOB563 -- Assignment #2'
author: "Dennis Psaroudakis"
date: "January 26th 2019"
output:
  pdf_document:
    dev: pdf
    latex_engine: lualatex
    md_extensions: +link_attributes+multiline_tables
    pandoc_args: --listings
mainfontoptions: Numbers=OldStyle
monofont: Linux Libertine Mono O
mainfont: Linux Libertine O
papersize: letter
sansfont: Linux Biolinum O
language: en-US
...

## 2