### 2019-03-12 assignments/5/Assignment_5.Rmd
Assignment 5.3 done

3
F84 is not available in raxml, so I'm using F81
Raxml: Final LogLikelihood -54568.308901
FastMe: Final LogLikelihood -54674.432495

4